---
title: "HW 8"
author: "Nitish Neelagiri"
date: "November 22, 2015"
output: pdf_document
---

#Question 1

##(a)
```{r}
setwd("~/")
library(survival)
survivalData <- read.csv("surv_times_data.csv")
attach(survivalData)

fit1_km = survfit(Surv(Y1, Delta1) ~ 1)
fit2_km = survfit(Surv(Y2, Delta2) ~ 1)

plot(fit1_km$time, fit1_km$surv, xlab = "time", ylab = expression(hat(S)), 
  xlim = range(c(Y1, Y2)), ylim = c(0, 1), type = "s", col = "blue", lwd = 2)
lines(fit2_km$time, fit2_km$surv, type = "s", col = "red", lwd = 2)
legend(2, 0.4, legend = c("Group 1", "Group 2"), lwd = c(2, 2), col = c("blue", "red"), 
  bty = "n")
```

From the output plot, we can see that the survival probability for Group 2 is higher than Group 1 at a particular time event. So, Group 2 has better survival prognosis.

##(b)
```{r}
fit1_km_emp = survfit(Surv(Y1, rep(1, each = length(Y1))) ~ 1)
fit2_km_emp = survfit(Surv(Y2, rep(1, each = length(Y1))) ~ 1)

plot(fit1_km_emp$time, fit1_km_emp$surv, xlab = "time", ylab = expression(hat(S)), xlim = range(c(Y1, Y2)), ylim = c(0, 1), type = "s", col = "blue", lwd = 2)
lines(fit2_km_emp$time, fit2_km_emp$surv, type = "s", col = "red", lwd = 2)
legend(2, 0.4, legend = c("Group 1", "Group 2"), lwd = c(2, 2), col = c("blue", "red"), 
  bty = "n")
```

We can observe from the plots that the empirical curve has low survival probability compared to Kaplan-Meier curve at a particular time event. This means that the inclusion of the censored data has increased the survival probability. Thus the censored data is more than the actual time event, meaning the right censored data. 

##(c)
```{r}
time_hat_Y1 = c(0, fit1_km$time)
surv_hat_Y1 = c(1, fit1_km$surv)

time_hat_Y2 = c(0, fit2_km$time)
surv_hat_Y2 = c(1, fit2_km$surv)

mu_hat_km_Y1 = 0
mu_hat_km_Y2 = 0

for(i in 2:length(surv_hat_Y1)) {
  mu_hat_km_Y1 = mu_hat_km_Y1 + surv_hat_Y1[i - 1] * (time_hat_Y1[i] - time_hat_Y1[i - 1])
}

for(i in 2:length(surv_hat_Y2)) {
  mu_hat_km_Y2 = mu_hat_km_Y2 + surv_hat_Y2[i - 1] * (time_hat_Y2[i] - time_hat_Y2[i - 1])
}

#Mean Survival Time for Group 1
sum((time_hat_Y1[2:101] - time_hat_Y1[1:100]) * surv_hat_Y1[1:100])

#Mean Survival Time for Group 2
sum((time_hat_Y2[2:101] - time_hat_Y2[1:100]) * surv_hat_Y2[1:100])
```

##(d)
```{r}
#Quantile for Group 1:
quantile(fit1_km, c(0.25, 0.5 ,0.75))

#Quantile for Group 2:
quantile(fit2_km, c(0.25,0.5,0.75))
```

##(e)
```{r}
Y <- c(Y1, Y2)
Delta <- c(Delta1, Delta2)
GRP = factor(rep(0:1, each = 183))
survdiff(Surv(Y, Delta) ~ GRP)
```
The chi squared value is 51.9 and p-value is 5.78e-13.

##(f)
```{r}

## General numerical optimization.
log_lik_Y = function(theta, Y, Delta) {
  mu = theta[1]
  sg = theta[2]

  obj_Y = sum(Delta * log(dnorm(Y, mu, sg))) + 
    sum((1 - Delta) * log(1 - pnorm(Y, mu, sg)))
  
  return(obj_Y)
}

neg_log_lik_Y = function(theta, Y, Delta) { return(-log_lik_Y(theta, Y, Delta)) }

theta_0_Y1 = c(mean(Y1), sqrt(var(Y1)))
theta_0_Y2 = c(mean(Y2), sqrt(var(Y2)))

theta_hat_Y1 = optim(theta_0_Y1, neg_log_lik_Y, Y = Y1, Delta = Delta1)
theta_hat_Y2 = optim(theta_0_Y2, neg_log_lik_Y, Y = Y2, Delta = Delta2)

#MLEs of Group1:
#Mean : 
theta_hat_Y1$par[1]
#SD:
theta_hat_Y1$par[2]

#MLEs of Group2:
#Mean:
theta_hat_Y2$par[1]
#SD:
theta_hat_Y2$par[2]

#Comparision of plots of Kaplan Meier and MLE based Survival curves. Red line represents MLE based survival curve.
#Group 1
tm_Y1 = seq(0, 24.3, length = 1000)
S_hat_Y1 = 1 - pnorm(tm_Y1, theta_hat_Y1$par[1], theta_hat_Y1$par[2])
plot(fit1_km)
lines(tm_Y1, S_hat_Y1, lwd = 2, col = "red")

#Group 2
tm_Y2 = seq(0, 28.8, length = 1000)
S_hat_Y2 = 1 - pnorm(tm_Y2, theta_hat_Y2$par[1], theta_hat_Y2$par[2])
plot(fit2_km)
lines(tm_Y2, S_hat_Y2, lwd = 2, col = "red")
```

(g)
```{r}
B <- 500
n <- nrow(survivalData)
delta_hat <- rep(NA,B)
for(b in 1:B){
  cat(".")
ii_1 <- sample(1:n, replace = TRUE)
ii_2 <- sample(1:n, replace = TRUE)
dta_b <- data.frame("Y1" = Y1[ii_1], "Y2" = Y2[ii_2], "Delta1" = Delta1[ii_1], "Delta2" = Delta2[ii_2])

theta_0_Y1_b <- with(dta_b, c(mean(Y1), sd(Y1)))
theta_0_Y2_b <- with(dta_b, c(mean(Y2), sd(Y2)))
fit_Y1_b <- with(dta_b, optim(theta_0_Y1_b, neg_log_lik_Y, Y = Y1, Delta = Delta1))
fit_Y2_b <- with(dta_b, optim(theta_0_Y2_b, neg_log_lik_Y, Y = Y2, Delta = Delta2))

delta_hat[b] = fit_Y1_b$par[1] - fit_Y2_b$par[1]
}

#95% Confidence Interval
quantile(delta_hat, c(0.025, 0.975))
```